---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Before getting started, make sure that you have installed the package:

```{r echo=T, eval=FALSE, message=FALSE}
devtools::install_github("setzler/DiDforBigData")
```

If the package is installed, make sure that you load it:

```{r echo=T, eval=T, message=FALSE}
library(DiDforBigData)
```


# 1. Choosing the Control Group

### The `control_group` argument

Callaway & Sant'Anna (2021) emphasize that, depending on empirical context, the parallel-trends assumption may only hold for one of these possible control groups:

- `"never-treated"`: These are the control units that are never observed receiving the treatment within the sampling time frame.
- `"future-treated"`: These are the control units that are observed receiving the treatment within the sampling time frame.
- `"all"`: These include both the "never-treated" and "future-treated" groups.

`DiDforBigData` makes it easy to choose among these possibilities by setting the `control_group` argument of the `DiD` function.

### Example

We start by simulating some data to work with:

```{r echo=T, eval=T, message=FALSE}
sim = SimDiD(seed=123, sample_size = 1000)
simdata = sim$simdata
print(simdata)
```


We can check what the true ATT is in the simulated data. In particular, let's focus on the ATT for Cohort 2007:

```{r echo=T, eval=T, message=FALSE}
print(sim$true_ATT[cohort==2007])
```


We also set up the variable names as a list so that `DiDforBigData` knows which variable is the outcome, which variable is the treatment cohort, etc.:

```{r echo=T, eval=T, message=FALSE}
varnames = list()
varnames$time_name = "year" 
varnames$outcome_name = "Y"
varnames$cohort_name = "cohort"
varnames$id_name = "id"
```


Now, we are ready to estimate DiD. We focus on the ATT for the 2007 cohort at event times -3,...,5. Initially, the control group is not set, so the function will use the default option  `control_group = "all"`:

```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, min_event = -3, max_event=5)

print(did$results_cohort[Cohort==2007])
```

We now consider changing the estimation to only use the "never-treated" control group:

```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, 
          control_group = "never-treated", min_event = -3, max_event=5)

print(did$results_cohort[Cohort==2007])
```

Note that the sample size for the control group, `Ncontrol`, is now substantially lower. Finally, we consider changing this to only use the "future-treated" group:

```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, 
          control_group = "future-treated", min_event = -3, max_event=5)

print(did$results_cohort[Cohort==2007])
```

We see that, because there are no future-treated cohorts left at event time +5, it is no longer possible to provide an ATT estimate at event time +5 if using the future-treated control group.

# 2. Avoiding anticipation


### The `baseperiod` argument 

In DiD designs, the researcher must choose a base pre-period against which differences are measured. A natural choice is `baseperiod = -1`, which means to use the time period just before treatment onset at event time 0. 

As discussed by Callaway & Sant'Anna (2021), one possibility is that the treatment group begins responding to treatment prior to treatment onset, which is called "anticipation." If anticipation begins at period -1, then differences measured relative to `baseperiod = -1` yield inconsistent DiD estimates. 

Fortunately, there is an easy solution: choose a base pre-period from before anticipation began. If it seems reasonable to assume that the treatment group could not have anticipated treatment 3 years before treatment onset, then `baseperiod = -3` should be free of anticipation.


### Example

We can simulate data that is subject to anticipation using the `anticipation` argument in the `SimDiD` function:

```{r echo=T, eval=T, message=FALSE}
sim = SimDiD(seed=123, sample_size = 200, anticipation = 2)
simdata = sim$simdata
print(simdata)
```


Let's focus on the average ATT across all cohorts. There are two periods of treatment effects prior to treatment, which we can verify by checking the true ATT from the simulation:

```{r echo=T, eval=T, message=FALSE}
print(sim$true_ATT[cohort=="Average"])
```

If we estimate DiD using the default argument that `baseperiod = -1`, the estimate will be biased and inconsistent:

```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, min_event = -3, max_event=3)

print(did$results_average)
```

We now set `baseperiod = -3` to avoid the anticipation at -1 and -2:

```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, 
          baseperiod = -3, min_event = -3, max_event=3)

print(did$results_average)
print(did$results_cohort)
print(sim$true_ATT)
```

We see that the estimate is now close to the true ATT.


# 3. Controlling for Time-varying Covariates


### The `covariate_names` entry


We sometimes worry that treatment cohorts are selected on time-varying observables, and those time-varying observables also directly affect the outcome. If the growth rate in the observables differs between the treatment and control groups, it creates a violation of the parallel-trends assumption:  the treatment and control groups would have experienced different growth profiles in the absence of treatment due to their different observables.

Fortunately, this is easy to fix: Since the confounding variables are observable, we just need to control for those observables. `DiDforBigData` requires only that the list of variable names, `varnames`, is modified to include a `covariate_names` entry. For example, `varnames$covariate_names = c("X1","X2")` tells it to control linearly for time-variation in X1 and X2.


### Example

There is an option in `SimDiD()` to add a couple of covariates. 

```{r echo=T, eval=T, message=FALSE}

sim = SimDiD(sample_size=1000, time_covars=TRUE)
simdata = sim$simdata
print(simdata)

print(sim$true_ATT[cohort==2007])
```

There are two covariates, X1 and X2. In particular, X2  differs in growth rates across treatment cohorts, which means that it causes violations of parallel trends if not controlled. We now verify that DiD gives the wrong answer for cohort 2007:

```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, min_event = -3, max_event=5)

print(did$results_cohort[Cohort==2007])
```


To control linearly for X1 and X2, we just need to add them to the `covariate_names` argument of `varnames`:

```{r echo=T, eval=T, message=FALSE}
varnames$covariate_names = c("X1","X2")
```

Now we check DiD with controls for time-variation in X1 and X2:


```{r echo=T, eval=T, message=FALSE}
did = DiD(inputdata = simdata, varnames = varnames, min_event = -3, max_event=5)

print(did$results_cohort[Cohort==2007])
```

We see that the bias has been removed thanks to the control variables.
